Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-25 TACAS2018/paper.tex
editing + comments for the cegis sec

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Synthesising dynamics}\label{sec:dynamic_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%\subsection{Using optimization refinement for CEGIS of control systems}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
In this section, we describe our technique for synthesizing safe
digital feedback controllers using CEGIS. Broadly speaking, CEGIS
denotes an iterative technique where each iteration performs inductive
generalisation based on counterexamples provided by a verification
engine.

We build upon this paradigm to create CEGIS-OR (CEGIS with Optimization
Refinement).
Our approach is illustrated in Figure~\ref{fig:CEGIS-OR} and consists of three main
phases:
  \note{explain broadly what each phase does}
\begin{itemize}
\item {\sc Abstract}: constructs of an unsound abstract model that over-approximates
the dynamics of the system.
\item {\sc synthesize}: generates a candidate controller and observer.  Since
  we are using coarse approximations, these do not necessarily meet any portion
  of the specification, but either through optimization or refinement become
  more precise (therefore more likely to meet the specification) with each iteration.
\item {\sc model}: refines or redefines the abstractions needed by {\sc Abstract}
  in order to reduce the probability of failure in the next iteration.
\item {\sc verify}: verifies the behaviour of the candidate controller and observer
  against a sound over-approximation of the concrete dynamics of the system for an infinite space over an infinite time horizon.
\end{itemize}

@@

The main advantage of this approach is two-fold:
\begin{itemize}
\item[1.] Optimizing the candidate solution in each iteration of the CEGIS loop
  enables us to obtain better candidates that are more likely to meet the
  specification or fail by a smaller margin.  This has the potential to reduce
  the number of iteration of the CEGIS loop required to obtain a solution.
\item[2.] We start with simpler (i.e. imprecise) abstractions that get
  refined into more complex ones as the procedure progresses. Thus, solutions
  that don't require a high level of precision are obtained fast. 
\end{itemize}

@@

We use two objective functions for optimization.  Both aim at reducing the
size of the reach space of the closed loop system while maintaining other
verification properties.  The first one aims at making the system as
convergent as possible.  This does not only ensure stability, but reduces
the response time of the system, moving away from the unsafe states faster.
The second one seeks to minimize rotations.  If the system rotates slower, it
is less likely to reach unsafe states before converging.
In both cases, the shape of the eigenspace may affect us negativey, so we
use a measure of the size of the reach space to guide our optimization.

@@

The CEGIS-OR loop for this approach is described next.
\paragraph{1. {\sc abstract}} creates an abstraction from the initial specificationas follows:
\begin{itemize}
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:observable}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{itemize}
\note{This paragraph needs to be moved before the itemize and it needs to be improved: give an intuition for what $\phi_{static}$ and $\phi_{synth}$ represent, e.g. the system's abstraction + stability spec + safety spec...}
There are two abstractions created by these specifications: $\phi_{synth}$ which is given to the synthesiser and $\phi_{static}$ which is given to the synthesis refinement loop. These abstractions will be refined independently, so we will not detail what each will contain at a given stage. We only note that at all times $\phi_{static} \models \phi_{synth}$.
%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy this tighter
specification, it feeds back to the {\sc Optimise} phase.
If the candidate is valid, we feed it to the next {\sc verify} stage, {\sc Discrete Time}.

>>>>

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Synthesising digital controllers with state observers}\label{sec:dynamic_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%\subsection{Using optimization refinement for CEGIS of control systems}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
In this section, we describe our technique for synthesizing safe
digital feedback controllers using CEGIS. Broadly speaking, CEGIS
denotes an iterative technique where each iteration performs inductive
generalisation based on counterexamples provided by a verification
engine.
%
We build upon this paradigm to create CEGIS-OR (CEGIS with Optimization
and Refinement).
Our approach is illustrated in Figure~\ref{fig:CEGIS-OR} and consists of four main
phases:
\begin{itemize}
\item {\sc Abstract}: constructs an unsound abstract model that over-approximates
the dynamics of the system.
\note{what does it mean to be over-approximating but unsound?}
\item {\sc synthesize}: generates a candidate controller and observer.
  As it makes use of the potentially coarse abstraction constructed by {\sc Abstract},
  the candidate doesn't necessarily meet the stability or safety specification.
  %% these do not necessarily meet any portion
  %% of the specification, but either through optimization or refinement become
  %% more precise (therefore more likely to meet the specification) with each iteration.
\item {\sc model}: refines the abstraction needed by {\sc Abstract}
  in order to reduce the probability of failure in the next iteration.
  \note{{\sc abstract} doesn't need an abstraction.}
  \note{how about optimization?}
\item {\sc verify}: verifies the behaviour of the candidate controller and observer
  against a sound over-approximation of the concrete dynamics of the system for an infinite space over an infinite time horizon. If the verification fails, produces a counterexample
  describing the reason for failure.
\end{itemize}

\note{this needs to be integrated in the above enumeration}
Our optimization procedure aims at reducing the
size of the reach space of the closed loop system while maintaining other
verification properties. \note{what are other verification properties?
if you mean safety and stability say so.}
%% The first one aims at making the system as
%% convergent as possible.  This does not only ensure stability, but reduces
%% the response time of the system, moving away from the unsafe states faster.
%% The second one seeks to minimize rotations.  If the system rotates slower, it
%% is less likely to reach unsafe states before converging.
%% In both cases, the shape of the eigenspace may affect us negativey, so we
%% use a measure of the size of the reach space to guide our optimization.

@@

The main advantage of this approach is two-fold:
\begin{itemize}
\item[1.] We start with simpler (i.e. imprecise) abstractions that get
  refined into more complex ones as the procedure progresses. Thus, solutions
  that don't require a high level of precision are obtained fast. 
\item[2.] Although {\sc synthesize} makes use of a potentially coarse abstraction,
  by optimizing the candidate solution in each iteration of the CEGIS loop
  we obtain better candidates that are more likely to meet the
  specification or produce high quality counterexamples in the
  {\sc verify} phase.
  This has the potential to reduce
  the number of iterations of the CEGIS loop required to obtain a solution.
\end{itemize}

@@

@@

\note{the next paragraph doesn't sound right, we need to separate the notions
  of abstraction and specification. they should not be the same. for now, just give an
intuition and we'll think more...}

The CEGIS-OR loop for this approach is described next.  \note{give an
  intuition for what $\phi_{static}$ and $\phi_{synth}$ represent,
  e.g. the system's abstraction + stability spec + safety spec...}
Note that we make use of two abstractions corresponding to two
specifications: $\phi_{synth}$ used by the synthesiser and
$\phi_{static}$ which is given to the synthesis refinement loop. These
abstractions will be refined independently, so we will not detail what
each will contain at a given stage. We only note that at all times
$\phi_{static} \models \phi_{synth}$.

\paragraph{1. {\sc abstract}} creates an abstraction from the initial specificationas follows:
\begin{itemize}
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:observable}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{itemize}
%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy this tighter
specification, it feeds back to the {\sc Optimise} phase.
If the candidate is valid, we feed it to the next {\sc verify} stage, {\sc Discrete Time}.
